Red-Giant Hot Jupiters: Brilliant Radio Beacons?

### 2015-08-26 ms.tex
Neil's section, 1 sentence per line.
further refinements to the sens and conf estimates

We consider in this section several current and near future radio observatories that 
operate at frequencies $\nu \lesssim 300~\rm MHz$ and can reach continuum 
sensitivities less than 1~mJy in a few hours.  The results are summarized
in Table~\ref{tab:sens}.
The primary limitations to detectability are noise, characterized by the sky
and system equivalent blackbody temperatures $T_{\rm sky}$ and $T_{\rm sys}$,
resolution, and source confusion.  
To estimate the source confusion $\sigma_c$ (in mJy beam$^{-1}$) reported in Table~\ref{tab:sens}, 
we use equations provided by \cite{condon+ransom2016}, reproduced below:
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!,~(\theta < 10\arcsec)
\end{eqnarray}
The confusion limit is dominated by active galactic nuclei, which have low polarization 
fractions $<2.5\%$ \citep{Stil2014} and slowly vary with time \citep{Condon1974,Condon2012}.
Therefore both the circularly polarized nature of the emission from hot Jupiters 
and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate 
the source confusion limit.

The LOw-Frequency ARray (LOFAR) operates in the 10-240~MHz range, split in two arrays
using a low band and high band antenna design (LOFAR-LBA and LOFAR-HBA), 
and different pad configuration \cite{vanHaarlem2013}.  
LOFAR's two most sensitive bands are centered at 60~MHz for LOFAR-LBA and 150~MHz 
for LOFAR-HBA respectively.  Given the strong frequency dependence of
the effective aperture of a dipole antenna, and that the system noise temperature 
$T_{\rm sys}$ for LOFAR is not fully determined by the sky noise $T_{\rm sky}$, 
we use the LOFAR Image Noise calculator\footnote{
Heald; \url{http://www.astron.nl/~heald/test/sens.php}} for simplicity, which 
is based on SKA Memo 113\footnote{
 Nijboer, Pandey-Pommier, \& de Bruyn; SKA Memo 113; 
\url{http://www.skatelescope.org/uploaded/59513\_113\_Memo\_Nijboer.pdf}}, for
our fiducial sensitivity and imaging parameters in Table~\ref{tab:sens}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htp]
\caption{Detectability of Hot Jupiters with Current and Future Radio Telescopes.}
\begin{center}
\begin{tabular}{c|ccccc} \hline \hline\\[-0.5pc]
Instrument & Band & Sensitivity & Time  & Resolution & Confusion limit\\
 & MHz & $\mu$Jy~hr$^{1/2}$ &  hr & \arcsec\ & $\mu$Jy bm$^{-1}$ \\[0.25pc] 
\hline\\[-0.5pc]
%ASKAP & 700-1700~MHz & 37 & 342.3 \\ % table 3 of http://www.atnf.csiro.au/projects/askap/config.html
GMRT  & 235~MHz &  165.6 & 6860 & 13 \\ % http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html
HERA  & 70~MHz  &  &  \\ % http://reionization.org/wp-content/uploads/2015/05/HERA0_imaging.pdf
% https://www.astron.nl/~heald/test/sens.php; 
% 24 core stations, 16 remote, 8 international, 488 subbands
LOFAR &  30-80  & 700 &  & 8.3 & 120 \\
LOFAR & 110-200 &  35 &  & 3.3 & 2.6\\  
LWA   &  30-80  &  &  \\
SKA   & 200  &  &  \\
UTR-2 & & & \\
VLA P-band & 270-470  & 138  & 4760 & \\
\hline
\end{tabular}
\end{center}
\label{tab:sens}
\tablecomments{Time estimates to reach 2~$\mu$Jy RMS.}
\end{table*}

>>>>

We consider in this section several current and near future radio observatories that 
operate at frequencies $\nu \lesssim 200~\rm MHz$ and can reach continuum 
sensitivities less than 1~mJy in a few hours.  The results are summarized
in Table~\ref{tab:sens}.
The primary limitations to detectability are noise, characterized by the sky
and system equivalent blackbody temperatures $T_{\rm sky}$ and $T_{\rm sys}$,
resolution, and source confusion.  
To estimate the source confusion $\sigma_c$ (in mJy beam$^{-1}$) reported in Table~\ref{tab:sens}, 
we use equations provided by \cite{condon+ransom2016}, reproduced below:
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!\!\!\!\!\!\!\!\!,~(\theta < 10\arcsec).
\end{eqnarray}
The source confusion limit is dominated by active galactic nuclei, which have low polarization 
fractions $\lesssim2.5\%$ \citep{Stil2014} and slowly vary with time \citep{Condon1974,Condon2012}.
Therefore both the circularly polarized nature of the emission from hot Jupiters 
and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate 
the source confusion limit.

The LOw-Frequency ARray (LOFAR) operates as two separate arrays defined by their 
high and low bands, differing configurations, and antenna designs.  
LOFAR-LBA operates in the 10-90~MHz range, and LOFAR-HBA operates in the 
100-240~MHz range \cite{vanHaarlem2013}.  We consider it as two separate
instruments here.
LOFAR-LBA's most sensitive band is centered at 60~MHz, and at 150~MHz 
for LOFAR-HBA.  Given the strong frequency dependence of
the effective aperture of a dipole antenna, and that the system noise temperature 
$T_{\rm sys}$ for LOFAR is not fully determined by the sky noise $T_{\rm sky}$, 
we use the LOFAR Image Noise calculator\footnote{
Heald; \url{http://www.astron.nl/~heald/test/sens.php}}, which 
is based on SKA Memo 113\footnote{
Nijboer, Pandey-Pommier, \& de Bruyn; 
\url{http://www.skatelescope.org/uploaded/59513\_113\_Memo\_Nijboer.pdf}}, for
our fiducial sensitivity and imaging parameters in Table~\ref{tab:sens}.

The Giant Metre-wave Radio Telescope (GMRT) is an array comprising thirty 
25-meter antennas, operating at frequencies 130-2000~MHz. The lowest
band (130-190~MHz) exhibits receiver noise comparable to the sky noise, and
hence we use their online calculator.\footnote{
\href{http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html}}

For the rest of the instruments considered, we assume $T_{\rm sys}\approx T_{\rm sky}$,
and use the fit from \cite{Rogers+Bowman2008} to data taken in the 100-200~MHz range:
\begin{equation}
T_{\rm sky} = T_{150} \left( \frac{\nu}{150~\rm MHz}\right)^{-\beta} + T_{\rm CMB} 
\end{equation}
where $T_{150} = 283.2~\rm K$ is the $T_{\rm sky}$ at 150~MHz, $\beta=2.47$,
and $T_{\rm CMB} = 2.73~\rm K$ is the contribution from the cosmic microwave 
background (CMB).
The RMS noise ${\sigma}$ in an interferometric observation, 
integration time $\tau$, bandwidth $\Delta\nu$, effective area $A_{\rm eff}$ 
($\approx0.7 \pi (D/2)^2$ for interferometers comprising antennas with dish size $D$),
and $N$ elements is \citep[see e.g.][]{condon+ransom2016} 
\begin{equation}
\sigma = \frac{2 k T_{\rm sys}}{A_{\rm eff} \, \sqrt{\Delta\nu \, \tau N (N-1)}} 
\left(\frac{\rm 10^{26} Jy}{\rm W m^{-2} Hz^{-1}}\right)
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htp]
\caption{Detectability of Hot Jupiters with Current and Future Radio Telescopes.}
\begin{center}
\begin{tabular}{c|ccccc} \hline \hline\\[-0.5pc]
Instrument & Band & Sensitivity & $\tau$  & Resolution & $5\sigma_c$\\
 & MHz & $\mu$Jy~hr$^{1/2}$ &  hr & \arcsec\ & $\mu$Jy bm$^{-1}$ \\[0.25pc] 
\hline\\[-0.5pc]
GMRT      & 130-190 & 430   &  460             &  20  & 1070 \\ 
HERA      &  30-80  & 540   &  730             & 880  & $1.4\times10^6$\\ 
LOFAR-LBA &  30-80  & 700   & 1225             & 8.3  & 120 \\ 
LOFAR-HBA & 110-200 &  35   &    3             & 3.3  & 2.6\\  
LWA       &  30-80  & 36000 &  3.2$\times10^6$ & 10.3 & 211\\
MWA       &  30-80  & 18000 &  8.1$\times10^5$ & 1030 & $1.4\times10^6$ \\
SKA-Low   & 50-350 &  3.5   &  0.03            & 7.9  & 57 \\
UTR-2 & & & \\
VLA P-band & 270-470  & 138  & 4760 & \\
\hline
\end{tabular}
\end{center}
\label{tab:sens}
\tablecomments{Time estimates to reach 20~$\mu$Jy RMS.}
\end{table*}

% HERA
% http://reionization.org/wp-content/uploads/2015/05/HERA0_imaging.pdf
%
% LOFAR
% https://www.astron.nl/~heald/test/sens.php; 
% 24 core stations, 16 remote, 8 international, 488 subbands
%
% GMRT
% http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html
%
